---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# theoddsapi

<!-- badges: start -->
<!-- badges: end -->

theoddsapi package provides a convenient wrapper for [the Odds API](https://the-odds-api.com/). The principal function `get_odds` provides a convenient, simple interface which allows you to leverage the API to pull odds from various different sports, books, and market types into tidy tibbles.

## Installation

Install the theoddsapi from GitHub with:

``` r
devtools::install_github("d-edison/theoddsapi")
```

## Configuration

In order to use this R package, you must first generate an API key (A free tier is available which grants 500 requests per month). To generate the key, navigate to [the Odds API's website](https://the-odds-api.com/) and click the "Get API Key" button. From there, you simply follow the instructions to get your key emailed to you. 

Once you have your key it is **highly recommended** to configure the R environment variable `THEODDS_API_KEY`. To do this, call either `Sys.setenv(THEODDS_API_KEY = "<your_key>")` or directly edit your `.Renviron` file.

## Usage

```{r attach}
library(theoddsapi)

# get all of the sports available
get_sports()

# find out how many requests are remaining
get_remaining_requests()

# make a request to get odds
get_odds("basketball_nba", mkt = "spreads")
```

